---
title: "Net size conversion factors"
output: html_document
editor_options: 
  chunk_output_type: console
---

Trying to create a model to account for undersampling of small zooplankters by CB nets
Using Bayesian models with brms to utilize complex model structures (hurdle_lognormal), but maybe there's an easier frequentist way to do this?
I Ran into computer issues with Stan and Rstudio repeatedly crashing, so I have not made much progress while I'm trying to figure that out.


```{r setup, include=FALSE}
require(tidyverse)
require(readxl)
require(lubridate)
require(brms)
```

# Load data and combine data

##Load data and rename columns

```{r}
EMP_CB <- read_excel("1972-2018CBMatrix.xlsx", 
                      sheet = "CB CPUE Matrix 1972-2018", 
                      col_types = c("numeric","numeric", "numeric", "numeric", "date", 
                                    "text", "text", "text", "numeric", 
                                    "text", rep("numeric", 62)))%>%
  rename(Volume=CBVolume)

EMP_Pump <- read_excel("1972-2018Pump Matrix.xlsx", 
                      sheet = "Pump CPUE Matrix 1972-2018", 
                      col_types = c("numeric","numeric", "numeric", "numeric", "date", 
                                    "text", "text", "text", "numeric", 
                                    "text", rep("numeric", 36)))%>%
  rename(Date=SampleDate, Volume=PumpVolume)
```

##Select columns in common, add sample and gear identifiers, combine data

```{r}
#Select columns in common between two datasets, rename taxa columns with distinct names

#Create vector of taxonomic variables
Taxvars<-c("LIMNOSPP", "LIMNOSINE", "LIMNOTET", "OITHDAV", "OITHSIM", "OITHSPP", "OTHCYCAD", "ALLCYCADULTS", "HARPACT", "CYCJUV", "LIMNOJUV", "OITHJUV", "OTHCYCJUV", "ALLCYCJUV", "COPNAUP", "EURYNAUP", "OTHCOPNAUP", "PDIAPNAUP", "SINONAUP", "ALLCOPNAUP", "ASPLANCH", "KERATELA", "OTHROT", "POLYARTH", "SYNCH", "SYNCHBIC", "TRICHO", "ALLROTIFERS", "BARNNAUP")

#create vector of taxonomic variables minus the summed "ALL" categories
Taxvars2<-c("LIMNOSPP", "LIMNOSINE", "LIMNOTET", "OITHDAV", "OITHSIM", "OITHSPP", "OTHCYCAD", "HARPACT", "CYCJUV", "LIMNOJUV", "OITHJUV", "OTHCYCJUV", "COPNAUP", "EURYNAUP", "OTHCOPNAUP", "PDIAPNAUP", "SINONAUP", "ASPLANCH", "KERATELA", "OTHROT", "POLYARTH", "SYNCH", "SYNCHBIC", "TRICHO", "BARNNAUP")

EMP_CB2<-EMP_CB%>%
  select(intersect(names(.), names(EMP_Pump)))%>% #select columns in common between two datasets
  select(-Survey, -SurveyCode, -SurveyRep, -EZStation, -Core)%>% #Remove some variables
  mutate(SampleID=paste(Date, Station))%>% #Create stationID identifier
  mutate(CPUE_Total=rowSums(select_at(., .vars=vars(Taxvars2))))%>% #Calculate total CPUE for total catch variable below
  rename_at(vars(Taxvars),
            ~paste0(., "_CB"))%>% #append "_CB" to taxonomic variables
  mutate_at(vars(paste0(Taxvars, "_CB")), list(l=~log(.+1)))%>% #create log(x+1) transformed counts
  mutate_at(vars(paste0(Taxvars, "_CB_l")), list(s=~(.-mean(.))/sd(.)))%>% #Standardize the logged counts to help models fit
  mutate(Catch_total_CB=CPUE_Total*Volume, #Add a total catch variable in case more zooplankton in a net clogs the net and helps catch more of the smaller zooplankters
         Catch_total_CB_s=(Catch_total_CB-mean(Catch_total_CB, na.rm=T))/sd(Catch_total_CB, na.rm=T))%>% #standardize total catch
  select(-Volume)

#Do the same for the Pump samples, except total pump catch shouldn't matter 
EMP_Pump2<-EMP_Pump%>%
  select(intersect(names(.), names(EMP_CB)))%>%
  select(-Survey, -SurveyCode, -SurveyRep, -EZStation, -Core)%>%
  mutate(SampleID=paste(Date, Station))%>%
  rename_at(vars(Taxvars),
              ~paste0(., "_Pump"))%>%
  mutate_at(vars(paste0(Taxvars, "_Pump")), list(l=~log(.+1)))%>%
  mutate_at(vars(paste0(Taxvars, "_Pump_l")), list(s=~(.-mean(.))/sd(.)))%>%
  select(-Volume)

#Join pump and CB data
EMP<-EMP_CB2%>%
  inner_join(EMP_Pump2, 
             by = c("Year", "Date", "Station", "Region", "Secchi", "Chl-a", "Temperature", "ECSurfacePreTow", "ECBottomPreTow", "SampleID"))%>%
  mutate(Secchi_s=(Secchi-mean(Secchi, na.rm=T))/sd(Secchi, na.rm=T),
         Chl_s=(`Chl-a`-mean(`Chl-a`, na.rm=T))/sd(`Chl-a`, na.rm=T),
         Temperature_s=(Temperature-mean(Temperature, na.rm=T))/sd(Temperature, na.rm=T))
```

# Graph

```{r}
ggplot(filter(EMP, Date>as_date("2007-01-01") & (LIMNOSINE_Pump>0 | LIMNOSINE_CB_l>0)), aes(x=LIMNOSINE_CB_l, y=LIMNOSINE_Pump_l))+
  geom_point()+
  theme_bw()+
  theme(panel.grid=element_blank())
```

# Model

##With Pump count as response variable 
```{r}
m<-censReg(log(LIMNOSINE_Pump+1) ~ log(LIMNOSINE_CB+1) * Secchi, data = filter(EMP, Date>as_date("2007-01-01")))

mbrm<-brm(LIMNOSINE_Pump~(LIMNOSINE_CB_l+Secchi_s+Chl_s+Temperature_s+Catch_total_CB_s)^3, 
          data = filter(EMP, Date>as_date("2007-01-01") & (LIMNOSINE_Pump>0 | LIMNOSINE_CB_l>0)), 
          family=hurdle_lognormal(),
          prior = prior(normal(0,5), class=b),
          chains=2, cores=2,
          control=list(adapt_delta=0.95))

mbrm2<-update(mbrm, formula.=.~.-Chl_s:Temperature_s:Catch_total_CB_s,
          chains=2, cores=2,
          control=list(adapt_delta=0.95))

mbrm3<-update(mbrm2, formula.=.~.-LIMNOSINE_CB_l:Chl_s:Catch_total_CB_s,
          chains=2, cores=2,
          control=list(adapt_delta=0.95))
```

##With CB count as the response variable, probably more appropriate since this is what we are modeling: how many small zooplankters are caught in the CB net?
```{r}
mbrmb<-brm(LIMNOSINE_CB~(LIMNOSINE_Pump_l+Secchi_s+Chl_s+Temperature_s+Catch_total_CB_s)^3, 
          data = filter(EMP, Date>as_date("2007-01-01") & (LIMNOSINE_Pump>0 | LIMNOSINE_CB_l>0)), 
          family=hurdle_lognormal(),
          prior = prior(normal(0,5), class=b),
          chains=2, cores=2,
          control=list(adapt_delta=0.95))
mbrmb<-add_criterion(mbrmb, c("loo", "waic"), reloo=T, cores=1)

mbrmb2<-update(mbrmb, formula.=.~.-LIMNOSINE_Pump_l:Secchi_s:Catch_total_CB_s -LIMNOSINE_Pump_l:Secchi_s:Temperature_s -Secchi_s:Chl_s:Temperature_s,
          chains=2, cores=2,
          control=list(adapt_delta=0.95))
mbrmb2<-add_criterion(mbrmb2, c("loo", "waic"), reloo=T)

mbrmb3<-update(mbrmb2, formula.=.~.-Secchi_s:Temperature_s:Catch_total_CB_s -LIMNOSINE_Pump_l:Secchi_s:Chl_s,
          chains=2, cores=2,
          control=list(adapt_delta=0.95))
mbrmb3<-add_criterion(mbrmb3, c("loo", "waic"), reloo=T)
```